---
title: "Conserved I-graph"
author: "Ellen Dimmen Chapple"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    number_sections: true
    toc_float: true
    theme: lumen
editor_options:
  chunk_output_type: console
---


```{r}

library(tidyverse)
library(igraph)
library(ggpattern)

```

# Finding cliques based on weight of edges

```{r}

# Read in expressologs
expressologs <- read.delim("Data/DATA/textFiles/all_expressed_genes.txt", sep = " ") %>% 
  filter(MaxpVal < 0.8)


expressologs_2 <- read.delim("Data/DATA/textFiles/all_expressed_genes.txt", sep = " ") %>% 
  filter(MaxpVal < 0.1)

cat("Orthogroups with at least one expressed pair:", expressologs %>% distinct(OrthologGroup) %>% nrow(), "\n")
cat("Orthogroups with at least one co-expressolog:", expressologs_2 %>% distinct(OrthologGroup) %>% nrow(), "\n")


```



```{r}

# 
load("Data/DATA/ones_and_zeros_all.RData") # See UPSET_main_ver4 rmd file!

df_with_sums <- ones_and_zeros_all %>%
  mutate(Angiosperms = rowSums(. [1:3])) %>%
  mutate(Cross = rowSums(. [4:12]))%>%
  mutate(Gymnosperms = rowSums(. [13:15])) %>%
  mutate(Conserved = rowSums(. [1:15]))

# Now we want to select only the ortholog groups with a gymnosperm sum of 3, while the others groups are 0.

conserved_ogs_gymno <- df_with_sums %>%
  filter(Angiosperms == 0 & Gymnosperms == 3 & Cross == 0)


conserved_ogs_gymno <- conserved_ogs_gymno %>% 
  filter(rownames(conserved_ogs_gymno) %in% expressologs$OrthologGroup)

cat("Number of gymnosperm-specific ortholog groups that have at least one co-expressed (not p-value cutoff) expressolog for all 3 pairs: ", nrow(conserved_ogs_gymno))

gymno_spec_OG <- rownames(conserved_ogs_gymno)

```




```{r}

# First - find all ortholog groups with at least one clique that involves only 3 species.
expressologs_from_max_cliques <- c()
max_clique_ortholog_groups <- c()
uuc_genes <- c()

for (g in gymno_spec_OG) {
  
  # g <- "OG0000773"
  
  expressologs_g <- expressologs %>% 
    filter(OrthologGroup == g) %>%
    mutate(UGeneSpecies1 = paste0(Species1, "-", GeneSpecies1),
           UGeneSpecies2 = paste0(Species2, "-", GeneSpecies2))
  
  nodes <- data.frame(name = unique(c(expressologs_g$UGeneSpecies1, expressologs_g$UGeneSpecies2)))
  
  edges <- data.frame(from = expressologs_g$UGeneSpecies1, 
                      to = expressologs_g$UGeneSpecies2,  weight = -log(expressologs_g$MaxpVal))
  
  # 
  net <- graph_from_data_frame(edges, directed = FALSE, vertices = nodes)
  
  max_clique <- largest_weighted_cliques(net)
  # Finds all of the maximal weighted cliques. 
  print(g)
  
  for( c in 1:length(max_clique)){
    
    # c <-1
    clique <- max_clique[[c]]
    clique <- attr(clique, "names")
    
    clique_species <- unique(str_split_fixed(clique, "-", n = 2)[,1])
    clique_genes <-  str_split_fixed(clique, "-", n = 2)[,2]

    if(length(unique(clique_species)) == 3 ) {
      
      max_clique_ortholog_groups <- c(max_clique_ortholog_groups, g)

      
      expressologs_in_clique <- expressologs_g %>%
        filter(UGeneSpecies1 %in% clique & UGeneSpecies2 %in% clique)%>% 
        mutate(cliqueSum = sum(MaxpVal))
      
      expressologs_from_max_cliques <- rbind(expressologs_from_max_cliques, expressologs_in_clique)
      
          
      uuc_genes <- rbind(uuc_genes, data.frame(
      Orthogroup = rep(g, length(clique_genes)),
      Genes = clique_genes,
      Species = clique_species
      
      ))

      
      
    }
  }
  
  
  
  
  
}
      


cat("Number of ortholog groups with at least one expressolog: ", length(unique(gymno_spec_OG)),"\n","Number of ortholog groups with at least one fully sized clique: ",length(unique(max_clique_ortholog_groups)))


```



# Creating data frame for group-wise counting


```{r}

# # Set p-value threshold
p_thr <- 0.1

max_clique_group_sum <- expressologs_from_max_cliques %>% 
  mutate(species_pair = paste0(Species1, Species2))%>%  
  group_by(OrthologGroup) %>% 
  arrange(cliqueSum) %>% 
  slice(1:3) %>% 
  ungroup()

speciesPairs <- tibble(SpeciesPair = max_clique_group_sum %>% distinct(species_pair) %>% pull(species_pair),
                       SpeciesPairClade = c( "Gymno","Gymno", "Gymno"))

#

max_clique_group_sum <-  left_join(max_clique_group_sum , speciesPairs, by = join_by("species_pair" == "SpeciesPair" ))


max_clique_group_sum <- max_clique_group_sum %>%
  group_by(OrthologGroup, species_pair) %>% 
  arrange(MaxpVal) %>%
  filter(MaxpVal < p_thr) %>% 
  group_by(OrthologGroup) %>% 
  mutate(conservedOG = n()) %>% 
  ungroup() %>% 
  pivot_wider(names_from = SpeciesPairClade, values_from = SpeciesPairClade) 
  

max_clique_group_sum$Gymno[is.na(max_clique_group_sum$Gymno)] <-0
max_clique_group_sum$Gymno[max_clique_group_sum$Gymno != 0] <-1

```


## Identifying best clique, and removing duplicated genes

```{r}



ultra_conserved_gymno <- max_clique_group_sum %>%
  mutate_at(c("Gymno"), as.numeric) %>%
  filter(conservedOG == 3 ) 

length(unique(ultra_conserved_gymno$OrthologGroup))

# 1) Identify the best clique per orthogroup

uuc_genes_filtered <- uuc_genes %>% 
  filter(Orthogroup %in% ultra_conserved_gymno$OrthologGroup)


length(unique(uuc_genes_filtered$Orthogroup))


uuc_genes_only_one_clique <- c()

for(i in unique(uuc_genes_filtered$Orthogroup)){
# i <- "OG0000078"
  print(i)
  
  lowest_scoring_clique <- ultra_conserved_gymno %>% 
    filter(OrthologGroup == i) 
  
  filtered_genes <- uuc_genes %>% 
    filter(Orthogroup == i) %>% 
    mutate(GeneSpecies = paste0(Species, "-", Genes)) %>% 
    filter(GeneSpecies %in% lowest_scoring_clique$UGeneSpecies1 | GeneSpecies %in% lowest_scoring_clique$UGeneSpecies2) %>% 
    distinct(GeneSpecies, .keep_all = T) %>% 
    select(-GeneSpecies)
  
  uuc_genes_only_one_clique <- rbind(uuc_genes_only_one_clique, filtered_genes)
  
}

# 2) Remove the duplicated genes
uuc_genes_only_one_clique <- uuc_genes_only_one_clique %>% 
  as.data.frame() %>% 
  group_by(Species) %>% 
  distinct(Genes, .keep_all = T) %>% 
  group_by(Orthogroup) %>% 
  mutate(sum = n()) %>% 
  filter(sum == 3) %>% 
  ungroup()

length(unique(uuc_genes_only_one_clique$Orthogroup))

# 3) Re-filter the uuc genes that now only contain one clique per orthogroup to also contain only complete cliques.

uuc_genes_only_one_clique <- uuc_genes_only_one_clique %>% 
  filter(Orthogroup %in% ultra_conserved_gymno$OrthologGroup)

length(unique(uuc_genes_only_one_clique$Orthogroup))


# PARTIAL CLIQUES

partial3_conserved_gymno <- max_clique_group_sum %>%
  mutate_at(c("Gymno"), as.numeric) %>%
  filter(conservedOG >=2)

length(unique(partial3_conserved_gymno$OrthologGroup))


# 1) Identify the best clique per orthogroup

part_genes_filtered <- uuc_genes %>% 
  filter(Orthogroup %in% partial3_conserved_gymno$OrthologGroup)

length(unique(part_genes_filtered$Orthogroup))

part_genes_only_one_clique <- c()

for(i in unique(part_genes_filtered$Orthogroup)){
   # i <- "OG0000033"
  print(i)
  
  lowest_scoring_clique <- partial3_conserved_gymno %>% 
    filter(OrthologGroup == i)
  
  filtered_genes <- uuc_genes %>% 
    filter(Orthogroup == i) %>% 
    filter(Genes %in% lowest_scoring_clique$GeneSpecies1 | Genes %in% lowest_scoring_clique$GeneSpecies2) %>% 
    distinct(Genes, .keep_all = T)
  
  part_genes_only_one_clique <- rbind(part_genes_only_one_clique, filtered_genes)
  
}

# 2) Remove the duplicated genes
part_genes_only_one_clique <- part_genes_only_one_clique %>% 
  group_by(Species) %>% 
  distinct(Genes, .keep_all = T) %>% 
  group_by(Orthogroup) %>% 
  mutate(sum = n()) %>% 
  filter(sum >= 2) %>% 
  ungroup()

length(unique(part_genes_only_one_clique$Orthogroup))

# 3) Re-filter the uuc genes that now only contain one clique per orthogroup to also contain only complete cliques.

part_genes_only_one_clique <- part_genes_only_one_clique %>% 
  filter(Orthogroup %in% partial3_conserved_gymno$OrthologGroup)


length(unique(part_genes_only_one_clique$Orthogroup))



```


## "Translating" gymnosperm genes to Aspen genes

Not possible to translate into Aspen genes as most of the orthologs unique for gymnosperms are not found within the other species. We can do it a little differently by using Norway Spruce genes. See separate GSEA-file!


```{r eval=FALSE, include=FALSE}
# Saving the uuc-genes


length(unique(uuc_genes_only_one_clique$Orthogroup))

save(uuc_genes_only_one_clique, file = "Data/DATA/uuc_orthologs_gymnosperms.RData")

uuc_genes_nor <- uuc_genes_only_one_clique %>% 
  filter(grepl("PICAB", Genes))

save(uuc_genes_nor, file ="Data/DATA/uuc_orthologs_gymnosperms_nor.RData")
# Now go to GSEA-file for gymnosperms!

```





