---
title: "Upset plot"
author: "Ellen Dimmen Chapple"
date: "2023-08-25"
output: 
  html_document:
    code_folding: hide
    toc: true
    number_sections: true
    toc_float: true
    theme: paper
editor_options:
  chunk_output_type: console
---



```{r}
library(dplyr)
library(tidyr)
library(gdata)
library(UpSetR)


```

```{r}
setwd("~/NMBU/M60-BIAS/EvoTree/githubEvoTree/Upset_plot")

```

Aim: create plot that shows the number of orthogroups that contains at least 1 gene from 1) all species, 2) all angiosperms, and 3) all gymnosperms.


# Loading files


```{r orthogroups and transcription data}
ortholog_group_file <- read.delim("Data/DATA/Orthogroups.100323.tsv", header = TRUE, sep = "\t")
species_list <- c("Asp", "Birch", "Cher", "Nor", "Scots", "Lodge")
angio_list <- c("Asp", "Birch", "Cher")
gymno_list <- c("Nor", "Scots", "Lodge")

```

# Extracting ortho groups present in all species

Each ortho group (OG) occurs only once, so the number of rows is the number of OGs which all species have in common!
In order to figure out the other species, simply select the species, and filter out the empty cells as done above.


```{r everbody}


ortho_general_filtering <- ortholog_group_file %>%
    mutate(Pinsyl.SHORT.cp.pep = Pinsyl.SHORT.pep) %>% 
    rename(
      Asp = Poptre.SHORT.pep,
      Birch = Betpen.SHORT.pep,
      Nor = Picabi.SHORT.pep,
      Scots = Pinsyl.SHORT.pep,
      Cher = Prunavi.SHORT.pep,
      Lodge = Pinsyl.SHORT.cp.pep,
      OrthoGroup = Orthogroup
    ) %>%
    mutate(Asp = gsub("Poptre_", "", Asp)) %>%
    mutate(Birch = gsub("Betpen_", "", Birch)) %>%
    mutate(Cher = gsub("Prunavi_", "", Cher)) %>% 
    select(OrthoGroup, species_list)



# 1) Identify all rows that contain at least 1 blank cell and add to a vector.

ortho_all <- ortho_general_filtering
rows_to_be_removed <- c()


for (i in 1:nrow(ortho_all)){  
  
  empty_row <- (ortho_all[i,] == "")
  sum <- sum(empty_row == T)

  if(sum >= 1) {
    rows_to_be_removed <- append(rows_to_be_removed, i, after = length(i))

     
  } else {  
    print(i)
    print (sum)
    } 
}
  
  


# 2) Remove the rows 
ortho_all <- ortho_all[-c(rows_to_be_removed),]


```


# Extracting ortho groups present only in angiosperms

```{r angios and gymnos }

# 1)  data frames 

ortho_angio <- ortho_general_filtering %>% 
  select("OrthoGroup",c(angio_list))

ortho_gymnos <- ortho_general_filtering %>% 
  select("OrthoGroup", c(gymno_list))

# 2) Create vectors with the row indices for locating empty cells

# GYMNOS


rows_to_be_removed <- c()
for (i in 1:nrow(ortho_gymnos)){  
  
  empty_row <- (ortho_gymnos[i,] == "")
  sum <- sum(empty_row == T)

  if(sum >= 1) {
    rows_to_be_removed <- append(rows_to_be_removed, i, after = length(i))

     
  } else {  
    print(i)
    print (sum)
    } 
}

removable_gymnos <- rows_to_be_removed


# ANGIOS
rows_to_be_removed <- c()
for (i in 1:nrow(ortho_angio)){  
  
  empty_row <- (ortho_angio[i,] == "")
  sum <- sum(empty_row == T)

  if(sum >= 1) {
    rows_to_be_removed <- append(rows_to_be_removed, i, after = length(i))

     
  } else {  
    print(i)
    print (sum)

    } 
}

removable_angios <- rows_to_be_removed

# 3) Filter the vectors so that each vector only contains a unique set of row indices. An overlap means that both groups contain at least one empty cell.


only_empty_in_gymnos <- removable_gymnos[!removable_angios %in% removable_gymnos]
only_empty_in_angios <- removable_angios[!removable_gymnos %in% removable_angios]

ortho_gymnos <- ortho_gymnos[c(only_empty_in_gymnos),]
ortho_angio <- ortho_angio[c(only_empty_in_angios),]



```



```{r}

paste0("Number of ortho groups shared by all sepcies: ", nrow(ortho_all))
paste0("Number of ortho groups unique to angiosperms: ", nrow(ortho_angio))
paste0("Number of ortho groups unique to gymnosperms: ", nrow(ortho_gymnos))

```


# Create the upset-plot

```{r}

input <- c(
  "Asp&Birch&Cher&Nor&Lodge&Scots" = nrow(ortho_all),
  "Asp&Birch&Cher" = nrow(ortho_angio),
  "Nor&Scots&Lodge" = nrow(ortho_gymnos)

)

upset_plot <- upset(fromExpression(input), 
      nintersects = 3, 
      nsets = 6, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.0, 
      line.size = 0.6
      )

upset_plot

```





- double check for neg correlations
- coexpressologs, pairwise

For 01.11.23:

- upset plot
- higher p-values were used
- har laget en legend reference for hm (mini-hm og et violinplot)


- mulig å få tilsendt en tidligere masteroppgave?


-> redo upset plot! See example in R. Redo data frame to 1's and 0's. but keep this one
make pairwise comparison, and then some criterea specific for each group (do this later)
















